---
title: "trachops_diet"
author: "may"
date: "9/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

QUESTIONS: 

Which sounds did bats approach or attack? 

Which had higher responses? 







```{r}
library(GmAMisc) 
library(readxl)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(RColorBrewer)
library(plotly)
library(car)
library(ordinal) 
library(emmeans)
library(multcompView)
```

```{r}

trac_sound<-read.table("https://raw.githubusercontent.com/maydixon/trachops_diet/master/Trachops_sound.09.09.19.txt", sep="\t", header=TRUE) #web
#View(trac_sound)



# Recode if.not.3 so that 0= no response, 1= ears twitch, 2=orient, 3=flight to, 4= lands
#Make Response, if.not.3, response.reviewed ordinal
#basically just have to make current 4 = 3 and 5 = 4

trac_sound<- trac_sound %>% mutate(if.not.3 = case_when(
      if.not.3 == "0" ~ "0" ,
      if.not.3 == "1" ~ "1",
      if.not.3 == "2" ~ "2",
      if.not.3 == "4" ~ "3",
      if.not.3 == "5" ~ "4"))

#Make a few columns factors

#trac_sound[ c("RESPONSE","RESPONSE_REVIEWED" ,"if.not.3")] <- lapply(trac_sound[ c("RESPONSE","RESPONSE_REVIEWED" ,"if.not.3")], as.factor)


# Remove unresponsive bat? 
trac_sound <-subset( trac_sound, trac_sound$Bat_ID !="Venom")


#relevel PLayback factors for plotting later
levels(trac_sound$PLYBAK)
trac_sound$PLYBAK <- factor(trac_sound$PLYBAK, levels = c("Eng_pus","Cra_fitz", "Sci_bou", "Pri_tae",   "Den_ebr", "Dia_dia",  "Flo_mel", "Ano_dry", "Ano_wet","Pur_ton"))
levels(trac_sound$PLYBAK)
```

Exploration
```{r}
str(trac_sound)
class(trac_sound$PLYBAK)
unique(trac_sound$PLYBAK)
class(trac_sound$NUMBER)
class(trac_sound$RESPONSE_REVIEWED)
unique(trac_sound$RESPONSE_REVIEWED)
```


```{r}
#hist(trac_sound$as.numeric(RESPONSE_REVIEWED)
```
Model of bat responses to the different sounds
```{r}
#https://stats.stackexchange.com/questions/368150/lsmeans-output-for-clmm-models-r
#https://cran.r-project.org/web/packages/emmeans/vignettes/sophisticated.html#ordinal


#model 

clmm_call_type<- clmm(as.factor(if.not.3) ~ PLYBAK + (1| Bat_ID), data= trac_sound, nAGQ = 10)
summary(clmm_call_type)
print(clmm_call_type)



#test for model proportional odds assumption (doesn't apply with 1 variable)
#scale_test(clm_call_type)


# Estimated marginal means: "mea

em_call_type <-emmeans(clmm_call_type, "PLYBAK", mode ="mean.class")

em_call_type
cld_call_type<- cld(em_call_type, letters=LETTERS)
cld_call_type
plot(cld_call_type)

x<- filter(trac_sound,PLYBAK =="Den_ebr" )
s= sum(as.numeric(x$if.not.3))/ length(as.numeric(x$if.not.3))
s

#
#ggsave("tracdiet_emmeans.jpg")


#Figure out the error
# Figure out how to make these more robust and what that means




```

Could potentially either bootstrap or permute this
Would have to randomize the data within groups (ID)
then run the CLMM
then find some measure for significance
extract that, see how often more significant than non-random model

or bootstrap


```{r}
summary(clmm_call_type)$coefficients

```

## PLOTTING

Make a doc with means and sd
And also full data in the background (greyed out)
```{r}

# Make response variable numeric
trac_sound$if.not.3 <- as.numeric(trac_sound$if.not.3)


# Mean response and SD per playback type
Resp_avg<-trac_sound %>% group_by(PLYBAK) %>%
      summarise(Resp_average = mean(if.not.3, na.rm=TRUE), sd = sd(if.not.3, na.rm = TRUE))

Resp_avg




```

some plot parameters
```{r}

#theme: 
TM <- theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), plot.background =element_blank(), legend.title = element_blank()) #tilted x-axis labels and title, no legend title

TM2<-theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.x=element_blank()) #tilted x-axis labels, no x title

TM2.5<-theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.x=element_blank(), legend.position="none") #tilted x-axis labels, no x title


TM4<-theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.x=element_blank(), legend.position="none") #tilted x-axis labels, no x title, no legend


TM1 <- theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), legend.title = element_blank(), plot.background =element_blank() ) #blank axes, blank legend title

TM3 <-  theme_calc() +theme(axis.title.x=element_blank(), axis.text.x = element_text(angle = 75, hjust=1), legend.position = "none") #tilted x-axis labels and title, for no size control, remoe base_size argument #for diagonal text, put angle = 45

TM4 <-  theme_calc() +theme(axis.title.x=element_blank(), axis.text.x=element_blank(), legend.position = "none") # no size control, no x axis labels

n_colors<- length(unique(trac_sound$PLYBAK))
PL <-  "Spectral"
PL <- colorRampPalette(brewer.pal(n_colors, "Set3"))
PL = colorRampPalette(brewer.pal(11, "Spectral"))

#palette with a lot of colors
c25 <- c("dodgerblue2","#E31A1C", # red
                "green4",
                "#6A3D9A", # purple
                "#FF7F00", # orange
                "black","gold1",
                "skyblue2","#FB9A99", # lt pink
                "palegreen2",
                "#CAB2D6", # lt purple
                "#FDBF6F", # lt orange
                "gray70", "khaki2",
                "maroon","orchid1","deeppink1","blue1","steelblue4",
                "darkturquoise","green1","yellow4","yellow3",
                "darkorange4","brown")
pie(rep(1,25), col=c25)


#this is the number of species in the plot/ dataset at any given time, can be used to change the palette

#+ theme(axis.text.x = element_text(angle = 45, hjust=1, size = 14), legend.position = "none") #format x-axis, and remove side legend #hjust 1 aligns labels
  #### settings for plot type 3####
  G3<- geom_point(position = position_jitter(width = 0.08), size = 5)
  G4<- geom_point( size = 5) #not jittered

#Y axis labels
  
Ex_labs <-  scale_x_discrete(breaks= c("Eng_pus","Cra_fitz", "Sci_bou", "Pri_tae",   "Den_ebr", "Dia_dia",  "Flo_mel", "Ano_dry", "Ano_wet","Pur_ton"),
     labels=c( c("E. pustulosus","C. fitzingeri", "S. boulengeri", "P. taeniatus",   "D. ebraccatus", "D. diastema",  "F. mellivora", "dry leaves - Anolis sp", "wet leaves - Anolis sp","Pure tone")))
  

  C3 <- coord_cartesian(ylim = c(0, 4))
  Plt2<-scale_colour_manual(values = PL(n_colors))
  Xlab2 <-  xlab("Acoustic stimuli")
  Ylab2 <-  ylab("Response score")
  Ylab3 <-  ylab(" Average response score")
  H2 <- 5  #height for plot output default
  W2<- 13 # height default
  
  # "#9E0142" "#CB334C" "#E95D47" "#F88D51" "#FDBE6E" "#FEE593" "#FFFFBF" "#EAF69E" "#BEE5A0" "#88CFA4" "#54AEAC" "#397EB8" "#5E4FA2"
  
   #Colors for graphs
cAJ<- scale_colour_manual(values = c("#9E0142")) #"#9E0142"
cGC<- scale_colour_manual(values = c("#CB334C"))
cLaB <- scale_colour_manual(values = c("#E95D47"))
```


```{r}
#plot mean responses and st. dev

Resp_avg

diet_plot <- ggplot() +
 geom_point(data = trac_sound, aes(x = PLYBAK, y = if.not.3, colour = PLYBAK), position = position_jitter(width = 0.08), size = 5) +
 C3 + TM3 + Xlab2 + Ylab2 + #all parameters for general data
geom_point(data = Resp_avg, mapping=aes(x = PLYBAK, y = Resp_average,    size = 4, position = position_jitter(width = 0.08))) + #parameters for the averages  #shape=17,color = PLYBAK,
geom_errorbar(data=Resp_avg, mapping=aes(x = PLYBAK,  ymin=Resp_average-sd, ymax=Resp_average+sd), width=.2)
diet_plot 
#mapping=aes(x = PLYBAK, y = Resp_average), 

Resp_avg

diet_plot <- ggplot(data = Resp_avg, aes(x = PLYBAK, y = Resp_average, colour = PLYBAK))
diet_plot <- diet_plot + G4+  C3 + TM3 + Xlab2 + Ex_labs + Ylab2 #all parameters for general data
diet_plot <- diet_plot + geom_errorbar(aes(ymin=Resp_average-sd, ymax=Resp_average+sd), width=.2) 

diet_plot <- diet_plot + geom_point(data = trac_sound, mapping=aes(x = PLYBAK, y = if.not.3), color = "light grey", alpha = 0.5, size = 4, position = position_jitter(width = 0.005)) #parameters for the averages  #shape=17,
class(trac_sound$if.not.3)
head(as.numeric)
diet_plot






#fix scale
#fix labels
#change colors

#rm(diet_plot)

```

