---
title: "trachops_diet"
author: "may"
date: "9/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

QUESTIONS: 

Which sounds did bats approach or attack? 

Which had higher responses? 







```{r}
library(GmAMisc) 
library(readxl)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(RColorBrewer)
library(plotly)
library(car)
library(ordinal) 
library(emmeans)
library(multcompView)
library(ggmosaic)
library(dplyr)
```

```{r}

trac_sound<-read.table("https://raw.githubusercontent.com/maydixon/trachops_diet/master/Trachops_sound.09.09.19.txt", sep="\t", header=TRUE) #web
#View(trac_sound)



# Recode if.not.3 so that 0= no response, 1= ears twitch, 2=orient, 3=flight to, 4= lands
#Make Response, if.not.3, response.reviewed ordinal
#basically just have to make current 4 = 3 and 5 = 4

trac_sound<- trac_sound %>% mutate(if.not.3 = case_when(
      if.not.3 == "0" ~ "0" ,
      if.not.3 == "1" ~ "1",
      if.not.3 == "2" ~ "2",
      if.not.3 == "4" ~ "3",
      if.not.3 == "5" ~ "4"))

#Make a few columns factors

#trac_sound[ c("RESPONSE","RESPONSE_REVIEWED" ,"if.not.3")] <- lapply(trac_sound[ c("RESPONSE","RESPONSE_REVIEWED" ,"if.not.3")], as.factor)

#make if.not.3 numeric 
trac_sound$if.not.3<- as.numeric(trac_sound$if.not.3)

# Remove unresponsive bat? 
trac_sound <-subset( trac_sound, trac_sound$Bat_ID !="Venom")


#relevel PLayback factors for plotting later: in order of amount found in diets


# (Pristimantis, Anolis, Craugastor, Engystymops, Dendropsophus, Scinax, Disporus, Florisuga, then tone)
levels(trac_sound$PLYBAK)

trac_sound$PLYBAK <- factor(trac_sound$PLYBAK, levels = c("Pri_tae", "Ano_dry", "Ano_wet", "Cra_fitz", "Eng_pus",  "Den_ebr", "Sci_bou",  "Dia_dia",  "Flo_mel", "Pur_ton"))
levels(trac_sound$PLYBAK)
```


Exploration
```{r}
str(trac_sound)

hist(trac_sound$RESPONSE_REVIEWED)

```


```{r}

clmm_call_type<- clmm(if.not.3 ~ PLYBAK + (1| Bat_ID), data= trac_sound_m, nAGQ = 10)
summary(clmm_call_type)
clmm_call_type_null<- clmm(if.not.3 ~ 1 + (1| Bat_ID), data= trac_sound_m, nAGQ = 10)
summary(clmm_call_type)
summary(clmm_call_type_null)

#compare the models
anova(clmm_call_type_null, clmm_call_type)

```




Model of bat responses to the different sounds
```{r}
#https://stats.stackexchange.com/questions/368150/lsmeans-output-for-clmm-models-r
#https://cran.r-project.org/web/packages/emmeans/vignettes/sophisticated.html#ordinal
trac_sound_m <- trac_sound
trac_sound_m$if.not.3 <- as.factor(trac_sound_m$if.not.3 )

#model 
# if.not.3 = response variable, ordinal scale 0-4 (Would 1-5 make the results more interpretable?)
# PLYBAK = the sound stimuli, IV, with 10 levels
# Bat_ID = Grouping/ random variable, 10 levels


clmm_call_type<- clmm(if.not.3 ~ PLYBAK + (1| Bat_ID), data= trac_sound_m, nAGQ = 10)
summary(clmm_call_type)




#test for model proportional odds assumption (doesn't apply with 1 variable)
#scale_test(clm_call_type)


# Estimated marginal means: "mean.class"
# "mean.class" obtains the average of these probability distributions as probabilities of the integers 1–5. Does this take into account random effects? 

em_call_type <-emmeans(clmm_call_type, "PLYBAK", mode ="mean.class")
em_call_type_latent <- emmeans(clmm_call_type, "PLYBAK")


em_call_type
em_call_type_latent 
cld_call_type<- cld(em_call_type, letters=LETTERS)
cld_call_type
plot(cld_call_type)

#how similar is this to just a straight mean? 
x<- filter(trac_sound,PLYBAK =="Den_ebr" )
s= sum(as.numeric(x$if.not.3))/ length(as.numeric(x$if.not.3))
s
# similar, except +1 because the scale here is automatically 1-5, rather than 0-4. How is this different? 

#
#ggsave("tracdiet_emmeans.jpg")

```

```{r}
#make emmeans output dataframe for ez plotting
em_data <- as.data.frame(cld_call_type)


# can plot this on the scale of my ordinal scale (0-4 rather than 1-5, which is default for mean.class, )
em_data.1<-em_data 

#subtract 1 from all estimates
em_data.1$mean.class<- em_data$mean.class -1
em_data.1$asymp.LCL<-em_data$asymp.LCL - 1
em_data.1$asymp.UCL<-em_data$asymp.UCL- 1

#write to CSV
em_data 
em_data.1
write.csv(em_data, file = "emm_trac_diet.csv")
write.csv(em_data.1, file = "emm.1_trac_diet.csv")
```

There aren't really good bootstrap or permutation solutions, and they would likely just show us th same thing. According to stats guy. 



## PLOTTING

Make a doc with means and sd
And also full data in the background (greyed out)
```{r}

# Make response variable numeric
#trac_sound$if.not.3 <- as.numeric(trac_sound$if.not.3)


# Mean response and SD per playback type
# (I call dply directly here bc some other package, probably ggplot, was calling plyr which made the group_by responses weird)
Resp_avg<-trac_sound %>% dplyr::group_by(PLYBAK) %>%
      dplyr::summarise(Resp_average = mean(if.not.3, na.rm=TRUE), sd = sd(if.not.3, na.rm = TRUE))

Resp_avg




```

some plot parameters
```{r}

#theme: 
TM <- theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), plot.background =element_blank(), legend.title = element_blank()) #tilted x-axis labels and title, no legend title

TM2<-theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.x=element_blank()) #tilted x-axis labels, no x title

TM2.5<-theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.x=element_blank(), legend.position="none") #tilted x-axis labels, no x title


TM4<-theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.x=element_blank(), legend.position="none") #tilted x-axis labels, no x title, no legend


TM1 <- theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), legend.title = element_blank(), plot.background =element_blank() ) #blank axes, blank legend title

TM3 <-  theme_calc() +theme(axis.title.x=element_blank(), axis.text.x = element_text(angle = 60, hjust=1), legend.position = "none") #tilted x-axis labels and title, for no size control, remoe base_size argument #for diagonal text, put angle = 45

TM3.5 <-  theme_calc() +theme(axis.title.x=element_blank(), axis.text.x = element_text(angle = 60, hjust=1), legend.position = "none") #tilted x-axis labels and title, for no size control, remoe base_size argument #for diagonal text, put angle = 45

TM4 <-  theme_calc() +theme(axis.title.x=element_blank(), axis.text.x=element_blank(), legend.position = "none") # no size control, no x axis labels

n_colors<- length(unique(trac_sound$PLYBAK))
PL <-  "Spectral"
PL <- colorRampPalette(brewer.pal(n_colors, "Set3"))
PL = colorRampPalette(brewer.pal(11, "Spectral"))

#palette with a lot of colors
c25 <- c("dodgerblue2","#E31A1C", # red
                "green4",
                "#6A3D9A", # purple
                "#FF7F00", # orange
                "black","gold1",
                "skyblue2","#FB9A99", # lt pink
                "palegreen2",
                "#CAB2D6", # lt purple
                "#FDBF6F", # lt orange
                "gray70", "khaki2",
                "maroon","orchid1","deeppink1","blue1","steelblue4",
                "darkturquoise","green1","yellow4","yellow3",
                "darkorange4","brown")
pie(rep(1,25), col=c25)


#this is the number of species in the plot/ dataset at any given time, can be used to change the palette

#+ theme(axis.text.x = element_text(angle = 45, hjust=1, size = 14), legend.position = "none") #format x-axis, and remove side legend #hjust 1 aligns labels
  #### settings for plot type 3####
  G3<- geom_point(position = position_jitter(width = 0.08), size = 5)
  G4<- geom_point( size = 5) #not jittered

#Y axis labels
  
Ex_labs <-  scale_x_discrete(breaks= c("Eng_pus","Cra_fitz", "Sci_bou", "Pri_tae",   "Den_ebr", "Dia_dia",  "Flo_mel", "Ano_dry", "Ano_wet","Pur_ton"),
     labels=c( c("E. pustulosus","C. fitzingeri", "S. boulengeri", "P. taeniatus",   "D. ebraccatus", "D. diastema",  "F. mellivora", "dry leaves - Anolis sp", "wet leaves - Anolis sp","Pure tone")))
  

  C3 <- coord_cartesian(ylim = c(0, 4))
  Plt2<-scale_colour_manual(values = PL(n_colors))
  Xlab2 <-  xlab("Acoustic stimuli")
  Ylab2 <-  ylab("Response score")
  Ylab3 <-  ylab(" Average response score")
  H2 <- 5  #height for plot output default
  W2<- 13 # height default
  
  # "#9E0142" "#CB334C" "#E95D47" "#F88D51" "#FDBE6E" "#FEE593" "#FFFFBF" "#EAF69E" "#BEE5A0" "#88CFA4" "#54AEAC" "#397EB8" "#5E4FA2"
  
   #Colors for graphs
cAJ<- scale_colour_manual(values = c("#9E0142")) #"#9E0142"
cGC<- scale_colour_manual(values = c("#CB334C"))
cLaB <- scale_colour_manual(values = c("#E95D47"))
```


```{r}
#plot mean responses and st. dev



diet_plot <- ggplot(data = Resp_avg, aes(x = PLYBAK, y = Resp_average, fill = "black"))
diet_plot <- diet_plot + G4+  C3 + TM3 + Xlab2 + Ex_labs + Ylab2 #all parameters for general data
diet_plot <- diet_plot + geom_errorbar(aes(ymin=Resp_average-sd, ymax=Resp_average+sd), width=.2) 

diet_plot <- diet_plot + geom_point(data = trac_sound, mapping=aes(x = PLYBAK, y = if.not.3), color = "light grey", alpha = 0.35, size = 4, position = position_jitter(width = 0.1, height = 0.000)) #parameters for the averages  #shape=17,
class(trac_sound$if.not.3)
head(as.numeric)
diet_plot


#ggsave("diet_plot_BW.pdf")
# I've just been moving points forward in inkscape, reversing order



#change colors

#rm(diet_plot)

```



## Or can do this is mosaic plot form

```{r}



mosaic_d2<- ggplot(data = trac_sound) +
geom_mosaic(aes(weight = 1, x = product(if.not.3, PLYBAK), fill=if.not.3),  na.rm=TRUE) +  
scale_fill_brewer(type="seq",  direction=1) +
Xlab2 + 
ylab("# of bats with response") +

#TM2
#scale_x_continuous(breaks= c("Eng_pus","Cra_fitz", "Sci_bou", "Pri_tae",   "Den_ebr", "Dia_dia",  "Flo_mel", "Ano_dry", "Ano_wet","Pur_ton"),labels=c( c("E. pustulosus","C. fitzingeri", "S. boulengeri", "P. taeniatus",   "D. ebraccatus", "D. diastema",  "F. mellivora", "dry leaves - Anolis sp", "wet leaves - Anolis sp","Pure tone"))) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),  axis.text.x = element_text( size = 12, angle = 45, hjust=1))+ 
      
guides(fill=guide_legend(title = "Response score"  , reverse = TRUE)) 
mosaic_d2 
# rm(mosaic_d2 ) 
# ggsave("mosaic_diet.jpg") 
 TM <- theme_calc(base_size = 25) + theme(axis.text.x = element_text(angle = 45, hjust=1), plot.background =element_blank(), legend.title = element_blank()) #tilted x-axis labels and title, no legend title


```

```{r}
#estra axis parameters on the right side
TM3.5 <-  theme_calc() +theme(axis.title.x=element_blank(), axis.text.x = element_text(angle = 60, hjust=1), legend.position = "none", axis.line.y.right = element_line(color = "grey"),  axis.ticks.y.right = element_line(color = "grey"), axis.text.y.right = element_blank(), axis.title.y.right = element_text(color = "grey"), axis.text.y.left = element_text(color = "black") ) #tilted x-axis labels and title, for no size control, remoe base_size argument #for diagonal text, put angle = 45

```



Same plot as above except with the model estimates and SD dev rather than the means and standard devs
```{r}
#em_data.1
#plot model estimates and 95 ci of estimates

diet_plot_em <- ggplot(data = em_data.1, aes(x = PLYBAK, y = mean.class, fill = "black"))
diet_plot_em <- diet_plot_em + G4+  C3 + TM3.5 + Xlab2 + Ex_labs +
  scale_y_continuous( "Mean probability of bat response score", sec.axis = sec_axis(~ . * 1, name = "Response score")) +

#ylab("Mean probability of bat response score") 

 geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), width=.2) 

diet_plot_em <- diet_plot_em + geom_point(data = trac_sound, mapping=aes(x = PLYBAK, y = if.not.3), color = "light grey", alpha = 0.35, size = 4, position = position_jitter(width = 0.1, height = 0))  

diet_plot_em 
diet_plot


#ggsave("diet_plot_em_BW.pdf")
# I've just been moving points forward in inkscape, reversing order



#change colors

#rm(diet_plot)

```
